---
title: "Enhanced Estimator"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Enhanced Estimator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(XSRecency)
```

In this vignette, we show how to use the enhanced estimator that incorporates prior HIV test results into the cross-sectional incidence estimate. In the process, we will show how to use the \code{\link{cephia}} data to estimate properties of recent infection testing algorithms.

### Estimate a phi function based on a recent infection testing algorithm from \code{\link{cephia}} data

#### Define algorithm of recent infection, and obtain data

Can either use the pre-loaded \code{\link{cephia}} data in this package,
or use your own that you have downloaded through the function \code{\link{downloadCephia}} with the optional argument `filepath`.

```{r}
# Define a recent infection as LAg <= 1.5 and viral load > 1000
f <- function(l, v){
 v <- ifelse(l > 1.5, 0, v)
 return(
   ifelse((l <= 1.5) & (v > 1000), 1, 0)
 )
}

# Get dataset
phidat <- createRitaCephia(assays=c("LAg-Sedia", "viral_load"), algorithm=f)

# Convert infection duration to years
phidat$ui <- phidat$ui / 365.25
head(phidat)
```

#### Estimate test-recent function

We can now use the CEPHIA data from above to estimate a phi function. We are doing this purely for the purposes of simulation. Usually, this is done internally in the \code{\link{estEnhanced}} function.

```{r}
# use this argument if you want to get an estimated phi function, rather than just MDRI summary
rita.props <- estRitaProperties(
  phidat=phidat,
  maxT=10,
  bigT=2,
  use_geese=TRUE, # need this when have multiple observations per individual
  formula="ri ~ poly(log(ui), degree=2, raw=T)",
  family=binomial(link="logit"),
  min_dt=TRUE, # need this when doing log(ui)
  return_all=T,
  plot_phi=T
)
```

### Simulate cross-sectional incidence data and prior HIV testing data

#### Cross-sectional incidence data

```{r}
# Use our estimated phi function to simulate recent infection indicators
sim <- simCrossSect(phi.func=rita.props$phi,
                    incidence_type="constant", prevalence=0.29, baseline_incidence=0.032)
set.seed(10)
df <- sim(5000)
```

#### Prior HIV testing data

Simulate prior HIV tests as 10% of HIV positive individuals having uniformly distributed tests over the last 4 years

```{r}
sim.pt <- simPriorTests(ptest.dist=function(u) runif(1, 0, 4),
                        ptest.prob=function(u) 0.1)

# Create prior testing data frame, only for those who are positive
ptdf <- sim.pt(df[df$di == 1,])
head(ptdf)
```

### Apply Enhanced Estimator

Apply the enhanced estimator using the CEPHIA data before that we used to simulate our current dataset,
and the same model for the test-recent function. See \code{\link{estEnhanced}} for details on additional values returned.

```{r}
estimate <- estEnhanced(
  n_p=nrow(ptdf),
  n=nrow(df),
  ptdf=ptdf,
  beta=0,
  beta_var=0,
  big_T=2,
  phidat=phidat,
  use_geese=TRUE,
  formula="ri ~ poly(log(ui), degree=2, raw=T)",
  family=binomial(link="logit"),
  min_dt=TRUE
)

estimate$est
estimate$var
```
## References

- Bannick, M. S., Donnell, D., Hayes, R., Laeyendecker, O., Gao F. (2023+). Enhanced Cross-Sectional HIV Incidence Estimators that Incorporate Prior HIV Test Results.
