---
title: "Snapshot and Adjusted Estimators"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Snapshot and Adjusted Estimators}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(XSRecency)
```

In this vignette, we show how to use the snapshot and adjusted estimators for cross-sectional incidence estimation. First, we will simulate cross-sectional data using functions from within this package. Then we illustrate how to use the estimation functions.

### Simulate Cross-Sectional Data

#### Define a test-recent function

```{r}
# Create an test recent function that approximates the following window and shadow period
# getTestRecentFunc() takes in window and shadow period in days as arguments
phi.func <- getTestRecentFunc(window=200, shadow=191)

ts <- seq(0, 10, by=0.01)
plot(phi.func(ts) ~ ts, type='l')

# Calculate the MDRI at 2 years for this function (we know that the window is 200/365.25)
window <- 200/365.25
mdri <- integratePhi(phi.func, maxT=2)
mdri
```

#### Simulate cross-sectional data based on constant incidence of 0.032 and prevalence of 29%.

```{r}
sim <- simCrossSect(phi.func=phi.func,
                    incidence_type="constant", prevalence=0.29, baseline_incidence=0.032)

set.seed(10)
data <- sim(5000)
head(data)
tail(data)
```

Records with \code{di == 0} are HIV-negative individuals, and \code{di == 1} are positive individuals.
The variables \code{ui} and \code{ri} are infection times (based on epidemiological parameters supplied) and recent infection indicators (based on the phi function), respectively.

### Apply Estimator

#### Snapshot Estimator

```{r}
estSnapshot(n_r=sum(data$ri, na.rm=T), n_n=sum(!data$di), n_p=sum(data$di), n=5000,
            mu=window, mu_var=0)
```

In some scenarios, not all individuals who test positive for HIV will receive a test. Here we show how to apply the estimator in these cases, using the argument `n_p_test`. The same argument can be used in the `estAdjusted` function.

```{r}
# Remove 50% of the recency tests to simulate what would happen if only 50% of
# HIV positive individuals received a recency test
data$mask <- rbinom(n=5000, size=1, prob=0.5)
data$ri_mask <- data$ri
data[data$mask == 1, "ri_mask"] <- NA

estSnapshot(n_r=sum(data$ri_mask, na.rm=T), n_n=sum(!data$di), n_p=sum(data$di), n=5000,
            mu=window, mu_var=0, n_p_test=sum(data$di & (!is.na(data$ri_mask))))
```

#### Adjusted Estimator

```{r}
# Suppose the FRR was known to be 0.1%
estAdjusted(n_r=sum(data$ri, na.rm=T), n_n=sum(!data$di), n_p=sum(data$di), n=100,
            omega=mdri, omega_var=0, beta=0.001, beta_var=0, big_T=2)
```

## References

- Kaplan, E. H., & Brookmeyer, R. (1999). Snapshot Estimators of Recent HIV Incidence Rates. *Operations Research*. 47(1): 29–37. [doi:10.1287/opre.47.1.29](https://doi.org/10.1287/opre.47.1.29)
- Kassanjee, R., McWalter, T. A., Bärnighausen, T., & Welte, A. (2012). A New General Biomarker-based Incidence Estimator. *Epidemiology*. 23(5): 721–728. [doi:10.1097/EDE.0b013e3182576c07](https://doi.org/10.1097/EDE.0b013e3182576c07)
